Delayed effects of climate on vital rates leads to demographic divergence in Amazonian forest fragments

### 2021-05-17 doc/paper.Rmd
re-order panels in size smooth plot

When summarizing across years and plots, the proportion of *Heliconia acuminata* that survived in CF and 1-ha was similarly high ($P_{surv} = 0.95$; Figure \@ref(fig:eda)b).
The proportion of surviving plants was lowest in the 2003--2004 transition year ($P_{surv} = 0.92$), which coincided with droughts in both the 2003 and 2004 rainy seasons (Figure \@ref(fig:eda)b,d) and was preceded by a drop in average plant size in the 2002--2003 transition year (Figure \@ref(fig:eda)a).
The lowest survival for 1-ha fragment plots ($P_{surv} = 0.93$) was for the 2005--2006 transition year, which encompassed a moderate drought in October 2005 and and wetter than average conditions (SPEI \> 0.5) in December 2005 and January 2006 (Figure \@ref(fig:eda)b,d).
Survival was size dependent in both continuous forest and 1-ha fragments ($p < 0.001$ for the effect of log-transformed plant size in year t on survival in year t+1 in both habitats).
While the survival probability of large plants approached 1 in both habitat types (Figure \@ref(fig:size-plot)a), the survival of the smallest plants was higher in 1-ha fragments.
However, the 95% confidence intervals for 1-ha fragments and continuous forest overlapped for all sizes.

Plants in continuous forest had an average of 2.9 shoots (± 1.8 SD) and were on average 40.6 cm tall (± 26.5 SD).
Plants in 1-ha fragments had on average \~`r format((1 - (2.5/2.9))*100, digits = 3)`% fewer shoots (2.5 ± 1.5 SD) and were \~`r format((1 - (36.2/40.6))*100, digits = 3)`% shorter (36.3 cm ± 24.1 SD).
Because the proxy for plant size used in our models was the product of these two metrics, plants in continuous forest were consequently substantially larger than those in forest fragments (150 ± 175 SD vs. 112 ± 141 SD, respectively).
<!--# Perhaps worth mentioning here that their distribution in different size classes is also different (skewed smaller) and likely why they are smaller on average). -->This disparity in plant size was most pronounced in the initial years of our surveys and then diminished over time (Figure \@ref(fig:eda)a), likely due to a combination of greater seedling recruitment in CF and plant mortality or shrinkage.
Mean plant size dropped dramatically in 2003 in both habitat types, corresponding with a severe drought during the February census (SPEI = -1.39) (Figure \@ref(fig:eda)d).
As with survival, size in year t was a significant predictor of size in year t+1 ($p < 0.001$ in both habitats).
While the effect was generally similar across size classes and habitat types, the impact of plant size on growth was greatest for mid-sized plants in continuous forest (Figure \@ref(fig:size-plot)b).

>>>>

When summarizing across years and plots, the proportion of *Heliconia acuminata* that survived in CF and 1-ha was similarly high ($P_{surv} = 0.95$; Figure \@ref(fig:eda)b).
The proportion of surviving plants was lowest in the 2003--2004 transition year ($P_{surv} = 0.92$), which coincided with droughts in both the 2003 and 2004 rainy seasons (Figure \@ref(fig:eda)b,d) and was preceded by a drop in average plant size in the 2002--2003 transition year (Figure \@ref(fig:eda)a).
The lowest survival for 1-ha fragment plots ($P_{surv} = 0.93$) was for the 2005--2006 transition year, which encompassed a moderate drought in October 2005 and and wetter than average conditions (SPEI \> 0.5) in December 2005 and January 2006 (Figure \@ref(fig:eda)b,d).
Survival was size dependent in both continuous forest and 1-ha fragments ($p < 0.001$ for the effect of log-transformed plant size in year t on survival in year t+1 in both habitats).
While the survival probability of large plants approached 1 in both habitat types (Figure \@ref(fig:size-plot)b), the survival of the smallest plants was higher in 1-ha fragments.
However, the 95% confidence intervals for 1-ha fragments and continuous forest overlapped for all sizes.

Plants in continuous forest had an average of 2.9 shoots (± 1.8 SD) and were on average 40.6 cm tall (± 26.5 SD).
Plants in 1-ha fragments had on average \~`r format((1 - (2.5/2.9))*100, digits = 3)`% fewer shoots (2.5 ± 1.5 SD) and were \~`r format((1 - (36.2/40.6))*100, digits = 3)`% shorter (36.3 cm ± 24.1 SD).
Because the proxy for plant size used in our models was the product of these two metrics, plants in continuous forest were consequently substantially larger than those in forest fragments (150 ± 175 SD vs. 112 ± 141 SD, respectively).
<!--# Perhaps worth mentioning here that their distribution in different size classes is also different (skewed smaller) and likely why they are smaller on average). -->This disparity in plant size was most pronounced in the initial years of our surveys and then diminished over time (Figure \@ref(fig:eda)a), likely due to a combination of greater seedling recruitment in CF and plant mortality or shrinkage.
Mean plant size dropped dramatically in 2003 in both habitat types, corresponding with a severe drought during the February census (SPEI = -1.39) (Figure \@ref(fig:eda)d).
As with survival, size in year t was a significant predictor of size in year t+1 ($p < 0.001$ in both habitats).
While the effect was generally similar across size classes and habitat types, the impact of plant size on growth was greatest for mid-sized plants in continuous forest (Figure \@ref(fig:size-plot)a).
